RDF Validation of Metadata \\ on Person-Level and Aggregated Data

### 2015-04-10 paper/paper7/paper7.tex
content-driven validation

\section{Introduction}
\tb{replace Data Cube by QB?}
%\tb{terminology: data in singular form / data set - NOT dataset / person-level data - NOT microdata / study - NOT survey}

@@

Bosch et al. identified in total \textcolor{red}{74} requirements to formulate RDF constraints; each of them corresponding to a constraint type. 
We published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} in which we explain each requirement (constraint type) in detail and give examples for each (represented by different constraint languages).
The knowledge representation formalism \emph{Description logics (DL)}, with its  well-studied theoretical properties, provides the foundational basis for each constraint type.
Therefore, this technical report contains mappings to DL to logically underpin each requirement and to determine which DL constructs are needed to express each constraint type \cite{BoschNolleAcarEckert2015}.
We recently published a technical report\footnote{\textcolor{red}{\label{technical-report-1}Available at: \url{http://arxiv.org/abs/XXXXX}}} (serving as first appendix of this paper) in which we describe constraints to validate metadata on person-level, aggregated data, and thesauri.
We assign each constraint to constraint types corresponding to RDF validation requirements or to data model specific constraint types\footnote{Requirements/Constraint types and constraints are uniquely identified by alphanumeric technical identifiers like \emph{R-1}}
\cite{BoschZapilkoWackerowEckert2015}.

@@

-----
%
%- different sets of constraints for different use cases / use case specific constraints
%
%- validation against minimal set of requirements / constraints
%
%-----

further ideas:

RDF validation scenarios require the closed-world assumption (CWA) (i.e., a statement is inferred to be false if it cannot be proved to be true).

use-case specific constraints, e.g., DCAT: searching for metadata
\tb{Thomas: ToDo}

@@

\subsubsection{Aggregated Metadata and Data}
\label{aggregated-metadata-and-data}

\tb{ToDo}

- for each dimension there should be a description and code lists.

- for each code list there should be a description.

- there should be a relationship to the underlying person-level data.

\subsubsection{Person-Level Metadata and Data}. 
\label{person-level-metadata-and-data}

@@

\textbf{Frequencies and Descriptive Statistics.}
The property \emph{disco:percentage} stands for the number of cases of a given code in relation to the total number of cases for a particular variable within a data set.
Percentage values are only valid when they are within the \emph{literal range} of 0 and 100 (\emph{R-45}; $\mathcal{SL}_{2}$).
{\em Mathematical Operations} (\emph{R-41, R-42}; e.g. date calculations and statistical computations like average, mean, and sum) are performed to ensure the integrity of data models.
The sum of percentage values of all codes of a variable code list, e.g., must exactly be 100 ($\mathcal{SL}_{2}$)
and the minimum of all variable codes do not have to be greater than the maximum ($\mathcal{SL}_{2}$).
Some constraints ensure the integrity of the data according to intended data model semantics (\emph{data model consistency}).
Codes (\emph{skos:Concept}), e.g, are ordered and therefore have fixed positions in an ordered collection (\emph{skos:OrderedCollection}) within variable representations.
In order to check the correctness of relative frequencies' calculations, the cumulative percentage (\emph{disco:cumulativePercentage}) of the current code must exactly be the cumulative percentage of the previous code (\emph{disco:cumulativePercentage})
plus the percentage value (\emph{disco:percentage}) of the current code ($\mathcal{SL}_{2}$).

@@

\textbf{Membership in Controlled Vocabularies.}
In many cases, resources must be members of listed controlled vocabularies (\emph{R-32}).
Types of summary statistics like minimum, maximum, and mode are maintained within a controlled vocabulary.  
Thus, summary statistics can only have \emph{disco:summaryStatisticType} relationships to \emph{skos:Concept}s which must be members of the controlled vocabulary \emph{ddicv:SummaryStatisticType}, a \emph{skos:ConceptScheme} (\emph{R-91: universal quantification}; $\mathcal{SL}_{2}$):
\begin{center}
\begin{DL}
SummaryStatistics $\sqsubseteq$ $\forall summaryStatisticType.A$ \\
$A \equiv Concept \sqcap \forall inScheme . B$ \\
$B \equiv ConceptScheme \sqcap \{SummaryStatisticType\}$
\end{DL}
\end{center}

@@

\textbf{Searching for (Meta)data.}

- DCAT

\textbf{Data Integration.}
use RDF validation for data integration

@@

\emph{Minimum/maximum/exact qualified cardinality restrictions} (\emph{R-74, R-75, R-76}) contain all those individuals that are connected by a property to at least/at most/exactly n different individuals/literals that are instances of particular classes or data ranges.
\emph{Minimum qualified cardinality restrictions} are defined for each property of multiple vocabularies like \emph{Disco}.
A \emph{disco:Questionnaire}, e.g., has at least one \emph{disco:question} relationship to \emph{disco:Questions} (\ms{Questionnaire $\sqsubseteq$ $\geq$1 question.Question}),
a \emph{disco:Variable} has at most one \emph{disco:concept} relationship to a theoretical concept (\emph{skos:Concept}) (\ms{Variable $\sqsubseteq$ $\leq$1 concept.Concept}), and every \emph{qb:DataSet} has (\emph{qb:structure}) exactly one associated \emph{qb:DataStructureDefinition} (\ms{DataSet $\sqsubseteq$ $\geq$1 structure.DataStructureDefinition $\sqcap$ $\leq$1 structure.DataStructureDefinition}).
%\footnote{\emph{DATA-CUBE-C-EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-02}}.
For particular data properties, values of predefined languages must be stated for determined number of times (\emph{R-48, R-49}).
Some controlled vocabularies contain literals in natural language, but without information what language has actually been used. Language tags might also not conform to language standards. 
Some concepts in a thesaurus are labeled in only one language, some in multiple languages. It may be desirable to have each concept labeled in each of the languages that also are used on the other concepts. This is not always possible, but incompleteness of language coverage for some concepts can indicate shortcomings of the vocabulary
(severity level: informational) \cite{MaderHaslhoferIsaac2012}.

>>>>

\section{Introduction}

%\tb{terminology: data in singular form / data set - NOT dataset / person-level data - NOT microdata / study - NOT survey}

@@

Bosch et al. identified in total 74 requirements to formulate RDF constraints; each of them corresponding to a constraint type. 
We published a technical report\footnote{Available at: \url{http://arxiv.org/abs/1501.03933}} in which we explain each requirement (constraint type) in detail and give examples for each (represented by different constraint languages).
The knowledge representation formalism \emph{Description logics (DL)}, with its  well-studied theoretical properties, provides the foundational basis for each constraint type.
Therefore, this technical report contains mappings to DL to logically underpin each requirement and to determine which DL constructs are needed to express each constraint type \cite{BoschNolleAcarEckert2015}.
We recently published a technical report\footnote{\textcolor{red}{\label{technical-report-1}Available at: \url{http://arxiv.org/abs/XXXXX}}} (serving as first appendix of this paper) in which we describe constraints to validate metadata on person-level, aggregated data, and thesauri.
We assign each constraint to constraint types corresponding to RDF validation requirements or to data model specific constraint types\footnote{Requirements/Constraint types and constraints are uniquely identified by alphanumeric technical identifiers like \emph{R-1}}
\cite{BoschZapilkoWackerowEckert2015}.

@@

%
%- different sets of constraints for different use cases / use case specific constraints
%
%- validation against minimal set of requirements / constraints
%
%-----

%further ideas:
%
%RDF validation scenarios require the closed-world assumption (CWA) (i.e., a statement is inferred to be false if it cannot be proved to be true).

%use-case specific constraints, e.g., DCAT: searching for metadata
%\tb{Thomas: ToDo}

@@

\textbf{Observations of Aggregated Data Sets.}
The purpose of some constraints is to ensure the integrity of
the data according to intended data model semantics (\emph{data model consistency}).
Every \emph{qb:Observation}, e.g., must have a value for each dimension
declared in its associated \emph{qb:DataStructureDefinition} ($\mathcal{SL}_{2}$)
and no two \emph{qb:Observations} in the same \emph{qb:DataSet}
may have the same value for all dimensions ($\mathcal{SL}_{1}$).
If a \emph{qb:DataSet} \emph{D} has a \emph{qb:Slice} \emph{S}, and \emph{S} has an
\emph{qb:Observation} \emph{O}, then the \emph{qb:DataSet} corresponding to \emph{O} must be \emph{D} ($\mathcal{SL}_{1}$).

%- for each dimension there should be a description and code lists.
%
%- for each code list there should be a description.
%
%- there should be a relationship to the underlying person-level data.

@@

\textbf{Descriptive Statistics.}
The property \emph{disco:percentage} stands for the number of cases of a given code in relation to the total number of cases for a particular variable within a data set.
Percentage values are only valid when they are within the \emph{literal range} of 0 and 100 (\emph{R-45}; $\mathcal{SL}_{2}$).
{\em Mathematical Operations} (\emph{R-41, R-42}; e.g. date calculations and statistical computations like average, mean, and sum) are performed to ensure the integrity of data models.
The sum of percentage values of all codes of a variable code list, e.g., must exactly be 100 ($\mathcal{SL}_{2}$)
and the minimum of all variable codes do not have to be greater than the maximum ($\mathcal{SL}_{2}$).
Codes (\emph{skos:Concept}), e.g, are ordered and therefore have fixed positions in an ordered collection (\emph{skos:OrderedCollection}) within variable representations.
In order to check the correctness of relative frequencies' calculations, the cumulative percentage (\emph{disco:cumulativePercentage}) of the current code must exactly be the cumulative percentage of the previous code (\emph{disco:cumulativePercentage})
plus the percentage value (\emph{disco:percentage}) of the current code (\emph{data model consistency}; $\mathcal{SL}_{2}$).

@@

\textbf{Membership in Controlled Vocabularies.}
In many cases, resources must be members of controlled vocabularies (\emph{R-32}).
Summary statistics types like minimum, maximum, and arithmetic mean, e.g., are maintained within a controlled vocabulary.  
Summary statistics can only have \emph{disco:summaryStatisticType} relationships to \emph{skos:Concept}s which must be members of the controlled vocabulary \emph{ddicv:SummaryStatisticType}, a \emph{skos:ConceptScheme} ($\mathcal{SL}_{2}$).
If a dimension property has a \emph{qb:codeList},
then the value of the dimension property on every \emph{qb:Observation} must be in the code list ($\mathcal{SL}_{2}$).

%\begin{center}
%\begin{DL}
%SummaryStatistics $\sqsubseteq$ $\forall summaryStatisticType.A$ \\
%$A \equiv Concept \sqcap \forall inScheme . B$ \\
%$B \equiv ConceptScheme \sqcap \{SummaryStatisticType\}$
%\end{DL}
%\end{center}

@@

%\textbf{Searching for (Meta)data.}
%
%- DCAT
%
%\textbf{Data Integration.}
%use RDF validation for data integration

@@

\emph{Minimum/maximum/exact qualified cardinality restrictions} (\emph{R-74, R-75, R-76}) contain all those individuals that are connected by a property to at least/at most/exactly n different individuals/literals that are instances of particular classes or data ranges.
\emph{Minimum qualified cardinality restrictions} are defined for each property of multiple vocabularies like \emph{Disco}.
A \emph{disco:Questionnaire}, e.g., has at least one \emph{disco:question} relationship to \emph{disco:Questions} (\ms{Questionnaire $\sqsubseteq$ $\geq$1 question.Question}),
a \emph{disco:Variable} has at most one \emph{disco:concept} relationship to a theoretical concept (\emph{skos:Concept}) (\ms{Variable $\sqsubseteq$ $\leq$1 concept.Concept}), and every \emph{qb:DataSet} has (\emph{qb:structure}) exactly one associated \emph{qb:DataStructureDefinition} (\ms{DataSet $\sqsubseteq$ $\geq$1 structure.DataStructureDefinition $\sqcap$ $\leq$1 structure.DataStructureDefinition}).

%DATA-CUBE-C-MINIMUM-QUALIFIED-CARDINALITY-RESTRICTIONS-
%02: Unique data set (IC-1 [3]) - Every qb:Observation has (qb:dataSet) ex-
%actly one associated qb:DataSet (Observation  ¥1 dataSet.DataSet [
%¤1 dataSet.DataSet).
%Severity level: ERROR
%
%(\emph{R-75: minimum qualified cardinality restrictions})
%
%DATA-CUBE-C-EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-
%02: Unique DSD (IC-2 [3]) - Every qb:DataSet has (qb:structure) exactly one
%associated qb:DataStructureDefinition (DataSet  ¥1 structure.DataStructureDefinition
%[ ¤1 structure.DataStructureDefinition).
%Severity level: ERROR
%
%(\emph{R-74: exact qualified cardinality restrictions})

%\footnote{\emph{DATA-CUBE-C-EXACT-QUALIFIED-CARDINALITY-RESTRICTIONS-02}}.
For particular data properties, values of predefined languages must be stated for determined number of times (\emph{R-48, R-49}).
Some controlled vocabularies contain literals in natural language, but without information what language has actually been used. Language tags might also not conform to language standards. 
Some concepts in a thesaurus are labeled in only one language, some in multiple languages. It may be desirable to have each concept labeled in each of the languages that also are used on the other concepts. This is not always possible, but incompleteness of language coverage for some concepts can indicate shortcomings of the vocabulary
(severity level: informational) \cite{MaderHaslhoferIsaac2012}.
